---
title: 'R Notebook: Warm up exercise with data frames'
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

This is a R Notebook that helps you to start working with data frames.

A data frame is a table or a two-dimensional array-like structure in which each column contains values of one variable and each row contains one set of values from each column.


## Fist step: Establish your working directory
Set your working directory with setwd ("P:/folder/folder") or go to Session menu-> Set Working Directory ->Choose Directory

```{r}
setwd("~/Desktop/Investigacion Operativa II/Machine Learning/R/T01 - Data Frames")
```


## Second step: Load the data
Load the file people-example.csv that needs to be saved in the working directory. You will use the function read.csv. Save the information in a variable called datos. This variable will be a data frame.

```{r}
people.example <- read.csv("~/Desktop/Investigacion Operativa II/Machine Learning/R/T01 - Data Frames/people-example.csv")
View(people.example)
```

The structure of the data frame can be seen by using str() function.


```{r}
str(people.example)
```

## Third step: Inspect the dataframe

Let's use some operations to check on the dataframe

Use the function head to see the first elements of the data frame

```{r} 
head(people.example)
```


Use the function tail to see the last elements of the data frame

```{r}
tail(people.example)
```


If you want to filter a specific column, for example Country column
```{r}
View(people.example$Country)
```

If you want to filter using more than one column, for example First.Name, Country and age columns
```{r}
df <- data.frame(people.example)
cols <- data.frame(df$First.Name,df$Country,df$age)
View(cols)
```

Calculate the max age value and its mean
```{r}
max(people.example$age)
mean(people.example$age)

```

Use summary function to summary data frame statistics
```{r}
summary(people.example)
```


# Operations with the dataframe itself
## Rename columns
Use rename function to change the names of First.Name and Last.Name columns. Use First and Last instead.
```{r}
# Load plyr package in order to use rename function
library(plyr)
library(magrittr)
# Add your code

names(df)[names(df) == "First.Name"] <- "First"
names(df)[names(df) == "Last.Name"] <- "Last"

colnames(df)

```

##Concatenate String in a new column
Add a new column in the data frame, called FullName, which will include the name and last name.

```{r}
# Add your code
FullName <- paste(df$First, df$Last)
newdf = cbind(df,FullName)
View(newdf)

```

## Removing columns
```{r}
# Will delete the column I just added
newdf$FullName <- NULL
View(newdf)
```
## Editing columns and programming
Making changes in the columns is usual in Data Science. Let's make a function to substract 50000 dollars in taxes to the ones who earns more than 300000 dollars.
There are two different ways of doing this: 1) using loops or 2) using apply family functions. Please, try both of them!

Option 1: Using loops



```{r}
for(i in length(df$Savings)){
  if(df[i,5] > 300000){
    df[i,5] = df[i,5] - 50000
  }
}
View(df[,5])


```



Option 2: Using apply family functions (see Quick R Tutorial)



```{r}
changeSavings<-function(saving){
  if(saving > 300000){
    saving = saving - 50000
  } else {
    saving = saving
  }
}

```

And now we are going to test our function before using it with the data frame
 
```{r}
a=changeSavings(5000000)
a
b=changeSavings(500)
b
```

Once we have tested the function works correctly, you will apply it to the data frame using sapply function.

```{r}
 #Add your code
lapply(df$Savings, changeSavings)
```



#Sorting values
Sort the values of the data frame by age column.

```{r}
#Add your code
ascending_age <- df[order(df$age),]

```


# Done!!
